---
title: "NYC_Crime"
author: "Jun Pu"
date: "2024-04-18"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Set up and Data Wrangling

```{r}
library(dplyr)
library(ggplot2)
library(shiny)
library(shinydashboard)
library(tidyr)
library(leaflet)
```


```{r}
# Read original file
crime.raw <- fread("crime.csv", stringsAsFactors = F)
crime.raw <- as.data.frame(crime.raw)
```

```{r}
#subset the data (taking columns needed for analysis)
crime=crime.raw %>%
  select(ARREST_DATE, KY_CD, OFNS_DESC, LAW_CAT_CD, ARREST_BORO, ARREST_PRECINCT, AGE_GROUP, PERP_SEX) %>%
  drop_na()
```

```{r}
#renaming the columns
crime=crime %>%
  rename(Date=ARREST_DATE, Code=KY_CD, Offense=OFNS_DESC,
         Type=LAW_CAT_CD, Boro=ARREST_BORO, Precinct=ARREST_PRECINCT, Age=AGE_GROUP, Sex=PERP_SEX)

#creating a new variable month
crime <- crime %>% 
  mutate(Month = format(as.Date(Date, format = "%m/%d/%Y"), "%m"))
```

```{r}
#save crime.csv for future use
write.csv(crime, file = "crime_nyc.csv")
```

# EDA and basic data description

```{r}
#basic data description
count_by_month=crime %>%
  group_by(Month) %>%
  count()

count_by_offense=crime %>%
  group_by(Offense) %>%
  count()

# Type: felony, misdemeanor, violation
count_by_type=crime %>%
  group_by(Type) %>%
  count()

count_by_boro=crime %>%
  group_by(Boro) %>%
  count()

count_by_precinct=crime %>%
  group_by(Precinct) %>%
  count()

count_by_age=crime %>%
  group_by(Age) %>%
  count()
```

```{r}
#add population column
## Source: https://www.citypopulation.de/en/usa/newyorkcity/
#crime$boro_year=paste0(crime$Boro,crime$Year)
population=function(x) {
  if (x=='B') return(1356476) # Bronx
  else if (x=='S') return(490687) #Staten Island
  else if (x=='K') return(2561225) #Brooklyn
  else if (x=='Q') return(2252196) #Queens
  else if (x=='M') return(1597451) #Manhattan
  else {
    return(0)
  }
}

crime$Population=sapply(crime$Boro, population)

```

```{r}
#further data clean up
crime$Month=as.numeric(crime$Month)


Month_converter=function(x) {
  if (x==1) return('Jan')
  else if (x==2) return('Feb')
  else if (x==3) return('Mar')
  else if (x==4) return('Apr')
  else if (x==5) return('May')
  else if (x==6) return('Jun')
  else if (x==7) return('Jul')
  else if (x==8) return('Aug')
  else if (x==9) return('Sep')
  else if (x==10) return('Oct')
  else if (x==11) return('Nov')
  else if (x==12) return('Dec')
}

crime$Month=sapply(crime$Month, Month_converter)

# Save this new dataset as crime_clean
write.csv(crime, file = "crime_clean.csv")

```

```{r}
crime_mo_bo=crime %>%
  group_by(Boro, Month) %>%
  summarise(n=n())

g=ggplot(crime_mo_bo, aes(x=Month, y=n, color=Boro))
g+geom_point() + theme(plot.subtitle = element_text(vjust = 1), 
    plot.caption = element_text(vjust = 1), 
    axis.text = element_text(size = 8, hjust = 1, 
        angle = 80))

```

## Inputs and Outputs

You can embed Shiny inputs and outputs in your document. Outputs are automatically updated whenever inputs change.  This demonstrates how a standard R plot can be made interactive by wrapping it in the Shiny `renderPlot` function. The `selectInput` and `sliderInput` functions create the input widgets used to drive the plot.

```{r eruptions, echo=FALSE}
inputPanel(
  selectInput("n_breaks", label = "Number of bins:",
              choices = c(10, 20, 35, 50), selected = 20),
  
  sliderInput("bw_adjust", label = "Bandwidth adjustment:",
              min = 0.2, max = 2, value = 1, step = 0.2)
)

renderPlot({
  hist(faithful$eruptions, probability = TRUE, breaks = as.numeric(input$n_breaks),
       xlab = "Duration (minutes)", main = "Geyser eruption duration")
  
  dens <- density(faithful$eruptions, adjust = input$bw_adjust)
  lines(dens, col = "blue")
})
```

## Embedded Application

It's also possible to embed an entire Shiny application within an R Markdown document using the `shinyAppDir` function. This example embeds a Shiny application located in another directory:

```{r tabsets, echo=FALSE}
shinyAppDir(
  system.file("examples/06_tabsets", package = "shiny"),
  options = list(
    width = "100%", height = 550
  )
)
```

Note the use of the `height` parameter to determine how much vertical space the embedded application should occupy.

You can also use the `shinyApp` function to define an application inline rather then in an external directory.

In all of R code chunks above the `echo = FALSE` attribute is used. This is to prevent the R code within the chunk from rendering in the document alongside the Shiny components.



